---
title: "Final Project"
author: "Nicholas Ilvento"
date: "4/1/2022"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(readxl)
EU_Accession_Data=read_excel("EU_Accession_Data.xlsx")
EU_Accession_Data_Predictions=read_excel("EU_Accession_Data_Predictions.xlsx")
```

```{r}
EU_Accession_Data$death_penalty=as.factor(EU_Accession_Data$death_penalty)
EU_Accession_Data$abortion=as.factor(EU_Accession_Data$abortion)
EU_Accession_Data$EFTA=as.factor(EU_Accession_Data$EFTA)
EU_Accession_Data$NATO=as.factor(EU_Accession_Data$NATO)
EU_Accession_Data$Yugoslavia=as.factor(EU_Accession_Data$Yugoslavia)
EU_Accession_Data$iron_curtain=as.factor(EU_Accession_Data$iron_curtain)
EU_Accession_Data$USSR=as.factor(EU_Accession_Data$USSR)
EU_Accession_Data$territory_disputes=as.factor(EU_Accession_Data$territory_disputes)
EU_Accession_Data$parliament=as.factor(EU_Accession_Data$parliament)
EU_Accession_Data$head_of_state=as.factor(EU_Accession_Data$head_of_state)
EU_Accession_Data$landlocked=as.factor(EU_Accession_Data$landlocked)
contrasts(EU_Accession_Data$death_penalty)
contrasts(EU_Accession_Data$abortion)
contrasts(EU_Accession_Data$EFTA)
contrasts(EU_Accession_Data$NATO)
contrasts(EU_Accession_Data$Yugoslavia)
contrasts(EU_Accession_Data$iron_curtain)
contrasts(EU_Accession_Data$USSR)
contrasts(EU_Accession_Data$territory_disputes)
contrasts(EU_Accession_Data$parliament)
contrasts(EU_Accession_Data$head_of_state)
contrasts(EU_Accession_Data$landlocked)
```
```{r}
cor(EU_Accession_Data[sapply(EU_Accession_Data,is.numeric)])
```

GDP_inflated_2022 and population_application 0.7727833
global_cities and GDP_inflated_2022 0.8461324
application_backlog and fertility_rate -0.70638301
members_at_application and fertility_rate -0.88318632
EU_Age and fertility_Rate -0.85119474

```{r}
plot(EU_Accession_Data$EU_age,EU_Accession_Data$length_months)
plot(EU_Accession_Data$members_at_application,EU_Accession_Data$length_months)
plot(EU_Accession_Data$application_backlog,EU_Accession_Data$length_months)
plot(EU_Accession_Data$GDP_application,EU_Accession_Data$length_months)
plot(EU_Accession_Data$GDP_inflated_2022,EU_Accession_Data$length_months) #looks like a relationship
plot(EU_Accession_Data$inflation_application,EU_Accession_Data$length_months) #looks like a relationship
plot(EU_Accession_Data$military_spending,EU_Accession_Data$length_months)
```

GDP_inflated_2022 and population_application 0.7727833
global_cities and GDP_inflated_2022 0.8461324
application_backlog and fertility_rate -0.70638301
members_at_application and fertility_rate -0.88318632
EU_Age and fertility_Rate -0.85119474

The country's GDP inflated to 2022 and the population is highly correlated at 0.7727833 and the GDP inflated to 2022 and the number of global cities are even more highly correlated at 0.8461324. I am getting rid of population and global cities since they do not appear to have any relationship with the number of months it takes to join the European Union but the inflated GDP appears to have a relationship of sorts. Fertility rate is highly negatively correlated with application backlog, members at application, and EU Age. Getting rid of fertility rate since the other three seem to have a better relationship with application duration.

```{r}
EU_Accession_Data$GDP_inflated_2022squared=(EU_Accession_Data$GDP_inflated_2022)^2
EU_Accession_Data$inflationsquared=(EU_Accession_Data$inflation_application)^2
EU_Accession_Data$members_at_applicationsquared=(EU_Accession_Data$members_at_application)^2
```

```{r}
cor(EU_Accession_Data$GDP_application,EU_Accession_Data$GDP_inflated_2022)
```

```{r}
testmodel1=lm(length_months~.-country-application-accession-GDP_inflated_2022squared-inflationsquared-members_at_application,data=EU_Accession_Data)
summary(testmodel1)
```

```{r}
testmodel2=lm(length_months~.-country-application-accession-GDP_application-inflationsquared-members_at_application,data=EU_Accession_Data)
summary(testmodel2)
```

```{r}
model=lm(length_months~.-country-application-accession-GDP_inflated_2022-GDP_application-inflationsquared-members_at_application,data=EU_Accession_Data)
summary(model)
```
Got rid of communism, it had a singularity with something.
GDP_inflated_2022 p-value = 0.362
GDP_application p-value = 0.582
GDP_inflated_2022squared p-value = 0.143
members_at_application p-value = 0.868
members_at_applicationsquared p-value = 0.841

```{r}
library(olsrr)
```

17! = 355,687,428,096,000
```{r,include=FALSE}
#allpossible=ols_step_all_possible(model)
#allpossible
``` 

```{r}
allpossible[which.min(abs(allpossible[,7])),]
```

```{r}
allpossiblemodel=lm(length_months~EU_age+abortion+EFTA+iron_curtain+GDP_inflated_2022squared,data=EU_Accession_Data)
summary(allpossiblemodel)
```

```{r}
library(MASS)
```

```{r}
stepwise=stepAIC(model,direction="both",trace=1)
summary(stepwise)
```
EU_age -- low p-value
inflation_application
military_spending
abortion -- lowest p-value
territory_dispute
iron_curtain -- low p-value
USSR
GDP_inflated_2022squared -- low p-value

```{r}
stepwisemodel=lm(length_months~EU_age+inflation_application+military_spending+abortion+territory_disputes+iron_curtain+USSR+GDP_inflated_2022squared,data=EU_Accession_Data)
summary(stepwisemodel)
```


```{r}
library(leaps)
```

```{r}
bestpossible=regsubsets(length_months~.-country-application-accession-GDP_application-GDP_inflated_2022-inflationsquared-members_at_application,data=EU_Accession_Data,nvmax=5,method="seqrep")
summary(bestpossible)
```
EFTA -- came up in two
head_of_state -- unique
EU_age -- came up in all three
GDP_inflated_2022squared -- came up in all three
inflation_application -- came up in two

```{r}
bestpossiblemodel=lm(length_months~EFTA+abortion+head_of_state+EU_age+GDP_inflated_2022squared+inflation_application,data=EU_Accession_Data)
summary(bestpossiblemodel)
```

Looks like as terms are added, the p-value, F-statistic, and adjusted R-squared value gets worse. I wanted to test just 5 variables but I also tested 4 and 6 just for comparison. All three of these models are horrible so ignore.

```{r}
plot(fitted(allpossiblemodel),resid(allpossiblemodel))
```

```{r}
plot(fitted(stepwisemodel),resid(stepwisemodel))
```

```{r}
sum(resid(allpossiblemodel))
```

```{r}
sum(resid(stepwisemodel))
```
The residual sum is closer to 0 on the stepwise model than the all-possible model.

```{r}
plot(stepwisemodel)
```

```{r}
summary(allpossiblemodel)
```

```{r}
summary(stepwisemodel)
```

Terms in Common between Stepwise and All-Possible Selection

EU_age - numerical - linear
abortion - categorical
iron_curtain - categorical
GDP_inflated_2022squared - numerical - quadratic


```{r}
combinedmodel=lm(length_months~EU_age+abortion+iron_curtain+GDP_inflated_2022squared,data=EU_Accession_Data)
summary(combinedmodel)
```
```{r}
plot(fitted(combinedmodel),resid(combinedmodel))
plot(EU_Accession_Data$EU_age,resid(combinedmodel))
plot(EU_Accession_Data$GDP_inflated_2022squared,resid(combinedmodel))
plot(EU_Accession_Data$GDP_inflated_2022squared,EU_Accession_Data$length_months)
plot(EU_Accession_Data$EU_age,EU_Accession_Data$length_months)
```


only use the variables in common between the all-possible and stepwise
look at two numerical variables from first all-possible model compared to residuals of the model
maybe transformations if not linear
then create model with transformed numerical and normal categorical variables

Questions 21 April 2022

I created the model of the variables in common between the stepwise and all-possible regression models
The F-statistic and the adjusted R-squared values are both lower than the stepwise model suggestion.
I created a residual plot for both numerical variables. Any suggestions?
I will create a model with the transformations from the residual plots.
If at the end of the day, the transformed model of this model with the in-common terms still has a smaller F-statistic and smaller adjusted R-squared than the stepwise, should I just use the suggested stepwise model instead?

To-Do List - actually really do this
1. Make single term model between y and x1 and x2 - graph against residuals to determine transformation
2. Do transformation on x-variables based off of residual plots
3. Check on interaction between variables - there is a special code for this (check notes) ***takes a long time***
4. Predictions


```{r}
modelx1=lm(length_months~EU_age,data=EU_Accession_Data)
plot(EU_Accession_Data$EU_age,resid(modelx1))
plot(fitted(modelx1),resid(modelx1))
```
Plot 1 - EU_age - between 0 and 300, between 300 and 410, between 411 and 541
Plot 2 - GDP_inflated_2022squared - between  0 and 
```{r}
library(dplyr)
```

```{r}
EU_Accession_Data=arrange(EU_Accession_Data,EU_age)
EU_Accession_Data$EU_ageW=rep(0,length(EU_Accession_Data$EU_age))
EU_Accession_Data$EU_age
```

```{r}
#Make sure to run above first
mean1=mean(EU_Accession_Data$EU_age[1:6])
sd1=sd(EU_Accession_Data$EU_age[1:6])
mean2=mean(EU_Accession_Data$EU_age[7:11])
sd2=sd(EU_Accession_Data$EU_age[7:11])
mean3=mean(EU_Accession_Data$EU_age[12:22])
sd3=sd(EU_Accession_Data$EU_age[12:22])
```

```{r}
#make sure to run previous two above first
(sd1^2)/mean1
(sd2^2)/mean2
(sd3^2)/mean3
(sd1^2)/(mean1^2)
(sd2^2)/(mean2^2)
(sd3^2)/(mean3^2)
(sd1^2)/sqrt(mean1)
(sd2^2)/sqrt(mean2)
(sd3^2)/sqrt(mean3)
```

Mean^2 has the most stable 3 values


```{r}
EU_Accession_Data$EU_ageW[1:6]=1/(mean1^2)
EU_Accession_Data$EU_ageW[7:11]=1/(mean2^2)
EU_Accession_Data$EU_ageW[12:22]=1/(mean3^2)
```

```{r}
EU_Accession_Data$EU_age_Weighted=((EU_Accession_Data$EU_age)*(EU_Accession_Data$EU_ageW))
```

```{r}
weightedmodelx1=lm(length_months~EU_age_Weighted,data=EU_Accession_Data)
summary(modelx1)
summary(weightedmodelx1)
plot(EU_Accession_Data$EU_age_Weighted,resid(weightedmodelx1))
```

```{r}
modelx2=lm(length_months~GDP_inflated_2022squared,data=EU_Accession_Data)
plot(EU_Accession_Data$GDP_inflated_2022squared,resid(modelx2))
plot(fitted(modelx2),resid(modelx2))
summary(modelx2)
```

```{r}
EU_Accession_Data=arrange(EU_Accession_Data,GDP_inflated_2022squared)
EU_Accession_Data$GDP_inflated_2022squared_W=rep(0,length(EU_Accession_Data$GDP_inflated_2022squared))
EU_Accession_Data$GDP_inflated_2022squared
```

```{r}
#Make sure to run above first
mean1=mean(EU_Accession_Data$EU_age[1:16])
sd1=sd(EU_Accession_Data$EU_age[1:16])
mean2=mean(EU_Accession_Data$EU_age[17:22])
sd2=sd(EU_Accession_Data$EU_age[17:22])
```

```{r}
#make sure to run previous two above first
(sd1^2)/mean1
(sd2^2)/mean2
(sd1^2)/(mean1^2)
(sd2^2)/(mean2^2)
(sd1^2)/sqrt(mean1)
(sd2^2)/sqrt(mean2)
```

Mean^2 has the most stable 2 values


```{r}
EU_Accession_Data$GDP_inflated_2022squared_W[1:16]=1/(mean1^2)
EU_Accession_Data$GDP_inflated_2022squared_W[17:22]=1/(mean2^2)
```

```{r}
EU_Accession_Data$GDP_inflated_2022squared_Weighted=((EU_Accession_Data$GDP_inflated_2022squared)*(EU_Accession_Data$GDP_inflated_2022squared_W))
```

```{r}
weightedmodelx2=lm(length_months~GDP_inflated_2022squared_Weighted,data=EU_Accession_Data)
summary(modelx2)
summary(weightedmodelx2)
plot(EU_Accession_Data$GDP_inflated_2022squared_Weighted,resid(weightedmodelx2))
```
```{r}
combinedmodelweights=lm(length_months~EU_age_Weighted+GDP_inflated_2022squared_Weighted+abortion+iron_curtain,data=EU_Accession_Data)
summary(combinedmodel)
summary(combinedmodelweights)
```

Weights made model worse. Forget weights.

```{r}
EU_Accession_Data$abortion_numerical=rep(0,length(EU_Accession_Data$abortion))
EU_Accession_Data$iron_curtain_numerical=rep(0,length(EU_Accession_Data$iron_curtain))
for(i in 1:length(EU_Accession_Data$abortion)){
  if(EU_Accession_Data$abortion[i]=="legal"){EU_Accession_Data$abortion_numerical[i]=1}
  else EU_Accession_Data$abortion_numerical[i]=0
}
for(i in 1:length(EU_Accession_Data$iron_curtain)){
  if(EU_Accession_Data$iron_curtain[i]=="yes"){EU_Accession_Data$iron_curtain_numerical[i]=1}
  else EU_Accession_Data$iron_curtain_numerical[i]=0
}
```

```{r}
cor((EU_Accession_Data$EU_age*EU_Accession_Data$GDP_inflated_2022squared),EU_Accession_Data$length_months)
cor((EU_Accession_Data$EU_age*EU_Accession_Data$abortion_numerical),EU_Accession_Data$length_months) #no
cor((EU_Accession_Data$EU_age*EU_Accession_Data$iron_curtain_numerical),EU_Accession_Data$length_months)
cor((EU_Accession_Data$GDP_inflated_2022squared*EU_Accession_Data$abortion_numerical),EU_Accession_Data$length_months)
cor((EU_Accession_Data$GDP_inflated_2022squared*EU_Accession_Data$iron_curtain_numerical),EU_Accession_Data$length_months) #no
cor((EU_Accession_Data$abortion_numerical*EU_Accession_Data$iron_curtain_numerical),EU_Accession_Data$length_months)
```

```{r}
cor((EU_Accession_Data$abortion_numerical*EU_Accession_Data$iron_curtain_numerical*EU_Accession_Data$EU_age),EU_Accession_Data$length_months)
cor((EU_Accession_Data$abortion_numerical*EU_Accession_Data$iron_curtain_numerical*EU_Accession_Data$GDP_inflated_2022squared),EU_Accession_Data$length_months)
cor((EU_Accession_Data$GDP_inflated_2022squared*EU_Accession_Data$iron_curtain_numerical*EU_Accession_Data$EU_age),EU_Accession_Data$length_months)
cor((EU_Accession_Data$GDP_inflated_2022squared*EU_Accession_Data$abortion_numerical*EU_Accession_Data$EU_age),EU_Accession_Data$length_months)
cor((EU_Accession_Data$GDP_inflated_2022squared*EU_Accession_Data$iron_curtain_numerical*EU_Accession_Data$EU_age*EU_Accession_Data$abortion_numerical),EU_Accession_Data$length_months)
```


EU_age - GDP_inflated_2022squared
EU_age - abortion
EU_age - iron_curtain
GDP_inflated_2022squared - abortion
GDP_inflated_2022squared - iron_curtain
abortion - iron_curtain

```{r}
interactionmodel1=lm(length_months~EU_age+GDP_inflated_2022squared+abortion+iron_curtain+EU_age*GDP_inflated_2022squared,data=EU_Accession_Data)
summary(interactionmodel1)
interactionmodel2=lm(length_months~EU_age+GDP_inflated_2022squared+abortion+iron_curtain+EU_age*GDP_inflated_2022squared*abortion,data=EU_Accession_Data)
summary(interactionmodel2)
interactionmodel2=lm(length_months~EU_age+GDP_inflated_2022squared+abortion+iron_curtain+EU_age*GDP_inflated_2022squared*abortion+EU_age*GDP_inflated_2022squared,data=EU_Accession_Data)
summary(interactionmodel2)
```

```{r}
combinedmodel=lm(length_months~EU_age+abortion+iron_curtain+GDP_inflated_2022squared,data=EU_Accession_Data)
summary(combinedmodel)
```

```{r}
summary(stepwisemodel)
```

```{r}
interactionmodelstep=lm(length_months~EU_age+inflation_application+military_spending+abortion+territory_disputes+iron_curtain+USSR+GDP_inflated_2022squared+EU_age*GDP_inflated_2022squared,data=EU_Accession_Data)
summary(interactionmodelstep)
```

```{r}
cor(EU_Accession_Data$EU_age*EU_Accession_Data$GDP_inflated_2022squared,EU_Accession_Data$length_months)
cor(EU_Accession_Data$EU_age*EU_Accession_Data$military_spending,EU_Accession_Data$length_months)
cor(EU_Accession_Data$EU_age*EU_Accession_Data$inflation_application,EU_Accession_Data$length_months)
cor(EU_Accession_Data$military_spending*EU_Accession_Data$GDP_inflated_2022squared,EU_Accession_Data$length_months)
cor(EU_Accession_Data$military_spending*EU_Accession_Data$inflation_application,EU_Accession_Data$length_months)
cor(EU_Accession_Data$inflation_application*EU_Accession_Data$GDP_inflated_2022squared,EU_Accession_Data$length_months)
```

```{r}
resid(stepwisemodel)
```


```{r}
plot(EU_Accession_Data$EU_age,resid(stepwisemodel))
plot(EU_Accession_Data$GDP_inflated_2022squared,resid(stepwisemodel))
plot(EU_Accession_Data$military_spending,resid(stepwisemodel))
plot(EU_Accession_Data$inflation_application,resid(stepwisemodel))
```


```{r}
summary(stepwisemodel)
```

```{r}
stepwisemodelfunction=function(EU_age,inflation_application,military_spending,abortion,territory_disputes,iron_curtain,USSR,GDP_inflated_2022squared) -5.804 + 0.4419*EU_age + 80.65*inflation_application - 1110*military_spending - 140.3*abortion + 46.51*territory_disputes + 69.4*iron_curtain - 39.8*USSR + 0.0000000000000000000001954*GDP_inflated_2022squared
```


```{r}
for(i in 1:length(EU_Accession_Data_Predictions$country)){
  EU_Accession_Data_Predictions$prediction[i]=stepwisemodelfunction(EU_Accession_Data_Predictions$EU_age[i],EU_Accession_Data_Predictions$inflation_application[i],EU_Accession_Data_Predictions$military_spending[i],EU_Accession_Data_Predictions$abortion[i],EU_Accession_Data_Predictions$territory_disputes[i],EU_Accession_Data_Predictions$iron_curtain[i],EU_Accession_Data_Predictions$USSR[i],EU_Accession_Data_Predictions$GDP_inflated_2022squared[i])
}
```


```{r}
EU_Predictions=data.frame(c(EU_Accession_Data_Predictions$country),c(EU_Accession_Data_Predictions$application),c(EU_Accession_Data_Predictions$prediction))
colnames(EU_Predictions)=c("country","application","predicted_duration")
EU_Predictions$rounded_prediction=round(EU_Predictions$predicted_duration,digits=0)
EU_Predictions
```

```{r}
library(lubridate)
```

```{r}
EU_Predictions$predicted_date=ymd(as.Date(EU_Predictions$application))%m+% months(EU_Predictions$rounded_prediction)
```

```{r}
EU_Predictions
```

```{r}
library(stats)
```

```{r}
interaction.plot(x.factor=EU_Accession_Data$EU_age,trace.factor=EU_Accession_Data$abortion,response=EU_Accession_Data$length_months)
```


```{r}
summary(stepwisemodel)
```



```{r}
interactionmodel=lm(length_months~EU_age*inflation_application*military_spending*abortion*territory_disputes*iron_curtain*USSR*GDP_inflated_2022squared,data=EU_Accession_Data)
summary(interactionmodel)
```


```{r}

```


